---
title: "DAV_Midterm_25thMarch"
author: "Akhil Patil"
date: "March 25, 2019"
output: word_document
---

DAV Mid - Term (Part 2)

```{r}
#install.packages("ISLR")
library(ISLR)
data(Auto)

```

A. Produce a scatterplot matrix which include all the variables in the data set.

```{r}
pairs(Auto)

```

B.Compute the matrix of correlations between the variables using the function cor(). You will need to exclude the "name" variable, which is qualitative.

```{r}
names(Auto)
cor(Auto[1:8])
```

C. perform a multiple linear regression with "mpg" as the response and all other variables except "name" as the predictors.Comment on the following:

```{r}
fit2 <- lm(mpg ~ . - name, data = Auto)
summary(fit2)

```


i. Is there a relationship between the predictors and the response ?

By testing the hypothesis   
                            h0:bi=0.
                            ha: bi != 0
Here h0 is Null Hypotheses ,ha is alternative hypotheses , bi are the predictor variables.

we get pvalue < 2.2e-16 which is less than signifance level of 0.01 1%. We can reject Null hypothesis. So , there is relationship between the mpg and other variables. 

ii. Which predictors appear to have a statistically significant relationship to the response ?

ANS :- when we check the p-values associated with each predictor variables t-statistic. We can confirm that except "cylinders", "horsepower" and "acceleration" other predictors like  "displacement" ,  "weight" , "year"  , "origin"  are  statistically significant.


iii. What does the coefficient for the "year" variable suggest ?

The variable coefficient of the "year" tells us that the average effect of an increase of 1 year is an increase of 0.7507727 in "mpg" and all other predictors remaining constant. That is cars become more fuel efficient every year by almost 1 mpg / year.

D. Obtain the 95% confidence interval of the coefficients

```{r}
confint(fit2,level = 0.95)

```


E. Produce diagnostic plots of the linear regression fit. Comment on any problems you see with the fit. Does the residual plots suggest any unusually large outliers ? Does the leverage plots identify any observations with unusually high leverages ?

```{r}
par(mfrow = c(2, 2))
plot(fit2)

```

Plot of residuals versus fitted values indicates the presence of a bit non linearity in the data. The plot of residuals versus leverage shows that there a few outliers (higher than 2 or lower than -2) and one high leverage point(point 14).


F)set aside one-forth of the observations to predict using the above linear model. compute the mean square error of the predicted values.

```{r}
auto_sub = Auto[sample(nrow(Auto), 294), ]
fit1=lm(mpg ~ . - name, data = auto_sub)
summary.fit1 = summary(fit1)
MeanSquareError = mean(summary.fit1$residuals^2)
MeanSquareError
```

