---
title: "NYC Jobs Analysis"
output: flexdashboard::flex_dashboard
---

Quantity Info
=====================================  
    
Column {data-width=600}
-------------------------------------
    
### Percentage Breakdown of Jobs by Agency
    
```{r}
# INCLUDE YOUR SETUP / DATASET IMPORT IN THIS CHUNK 

library(readr)
library(plotly)

NYC_Jobs <- read_csv("D:/Downloads/NYC_Jobs.csv")
#View(NYC_Jobs)

# CREATE A COPY OF YOUR DATASET HERE AND CHANGE COPY TO WHATEVER YOU LIKE
copy<- NYC_Jobs

# Q4."Quantity Info".1 BELOW LINE IS TO FIGURE OUT THE UNIQUE AGENCIES IN THE DATASET. PLEASE FILL OUT RELEVANT INFORMATION. CHANGE AGENCIES.VECTOR IF YOU WOULD LKE
agencies.vector <- unique(copy$Agency)
```

```{r}
# Q4."Quantity Info".2 BELOW WILL HANDLE THE SET UP FOR THE FOR LOOP. INITIALIZE THE VECTORS AS EMPTY VECTORS
agency.names = {}
agency.values = {}
index = 1

#Q4."Quantity Info".3 THIS IS THE FOR-LOOP. FIGURE OUT HOW TO ITERATE THROUGH AGENCIES.VECTOR
for(i in agencies.vector){
  manipulated.string<-i
  # Q4."Quantity Info".3.a.i.1 WE ARE CHECKING FOR THE "DEPARTMENT OF" OR "DEPT OF" STRINGS IN THE NAMES AND THEN DELETING THEM ACCORDINGLY
  
  # CHECK BELOW FOR "DEPARTMENT OF" in manipulated.string using grepl ()
  if(grepl("DEPARTMENT OF",manipulated.string)){
    # REPLACE "DEPARTMENT OF" WITH AN EMPTY STRING USING sub()
    manipulated.string<- sub("DEPARTMENT OF","",manipulated.string)
  } else if (grepl("DEPT OF",manipulated.string)) { # CHECK BELOW FOR "DEPT OF" in manipulated.string using grepl
     # REPLACE "DEPARTMENT OF" WITH AN EMPTY STRING USING sub()
    manipulated.string<- sub("DEPT OF","",manipulated.string)
  }
  
  # Q4."Quantity Info".3.a.ii.1 NOW WE WANT TO ADD THE MANIPULATED STRING TO OUR agency.names VECTOR. FIGURE OUT APPROPRIATE INDEX
  agency.names[index] = manipulated.string
  
  # Q4."Quantity Info".3.a.ii.2 WE ALSO WANT TO ADD THE NUMBER OF JOBS THAT EACH AGENCY POSTS IN THE ASSOCIATED agency.values VECTOR. which() will be useful here!
  agency.values[index] = length(which(NYC_Jobs$Agency ==  manipulated.string))
  
  # PLEASE DON'T CHANGE THIS VALUE. THIS IS NECESSARY. IN A COMMENT BELOW, COMMENT WHY
  # YOUR ANSWER:
  index = index + 1
}

#Q4."Quantity Info".3.b CREATING A DATAFRAME IN DESCENDING ORDER. NAME THE VARIABLES AND ORDER THE VECTOR 
merged.agency.vectors <- data.frame(agency.names,agency.values)

order.merged <- merged.agency.vectors[order(-agency.values),]

# just for fun, try plotting here and look at what goes wrong
library(plotly)

p <- plot_ly(mergedataset, labels = ~agency.names, values = ~agency.values, type = 'pie') %>%
  layout(title = 'Dissemination of Jobs(By Domain)',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

p

  
#Q4."Quantity Info".3.d THIS IS WHERE YOU'RE GENERATING YOUR CUTOFF. USE WHICH TO APPROACH THIS QUESTION
greater.values<- subset(order.merged, order.merged$agency.values >= 20) 
lower.values <- subset(order.merged, order.merged$agency.values < 20) 

#Q4."Quantity Info".3.d.i FIND THE SUM OF THE JOBS THAT ARE BELOW OUR THRESHOLD
other.sum <- sum(lower.values$agency.values)

# Q4."Quantity Info".3.d.i NOW WE WANT TO ADD AN "OTHER" VALUE TO OUR order.merged DATA FRAME DON'T TOUCH THIS CODE
add.other<-data.frame("OTHER", other.sum)
names(add.other) <- names(greater.values)
final.pie.chart <- rbind(greater.values, add.other)

# Q4."Quantity Info".3.e PLOT A PIE CHART LOOKING AT THE final.pie.chart$AGENCY.NAMES AND final.pie.chart$AGENCY.VALUES  USING PLOTLY. GIVE AN APPROPRIATE TITLE AND LAYOUT
jobs <-  plot_ly(final.pie.chart, labels = ~agency.names, values = ~agency.values, type = 'pie') %>%
  layout(title = 'Distribution ofthe Job offerings by Agency ',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
jobs
```

   
Column {data-width=400}
-------------------------------------
   
### Top 10 Job Agencies Offering Most # of Jobs

```{r}
library(gridExtra)
library(grid)

#c USING THE LINK FROM THE PDF, OUTPUT THE TOP 10 AGENCIES THAT OFFER THE MOST NUMBER OF JOBS

grid.table(final.pie.chart[1:10,])
final.pie.chart[1:10,]

```   
 

Page 2 {data-orientation=rows}
=====================================     
   
Row {data-height=600}
-------------------------------------

### Chart 1


```{r}
# Q4."Salary Breakdown".2 THIS IS THE SETUP FOR THE FOLLOWING FOR LOOP. 
library(plotly)
# positions should reference unique "Civil Service Title" values
positions<- unique(NYC_Jobs$`Civil Service Title`)
# initialize all of these to empty vectors
offered.jobs = {}
offered.values  = {}
unique.agencies = {}
average.salary = {}

index = 1
```

```{r}
# Q4."Salary Breakdown".3 THIS IS THE FOR LOOP. 
# choose iterator name for the for-loop
for(i in positions){
  # Q4."Salary Breakdown".3.a.i replace <value measuring number of these jobs offered> with any name
  #value.measuring.number.of.these.jobs.offered = {}
  # Q4."Salary Breakdown".3.a.i offered.jobs contains the names of the different jobs. At the appropriate index, input the iterator, as this is a reference to the name of the job
  offered.jobs[index] = i
  #  Q4."Salary Breakdown".3.a.ii offered.values contains the total number of jobs for this position being offered
  offered.values[index]  = length(which(NYC_Jobs$`Civil Service Title` == i))
  #  Q4."Salary Breakdown".3.a.iii unique.agencies contains the total number of unique agencies offering this job
  unique.agencies[index] =   length(unique(NYC_Jobs$Agency[which(NYC_Jobs$`Civil Service Title`== i)]))

  #  Q4."Salary Breakdown".3.a.iv average.salary contains the average salaries for these positions. ANALYZE THE DATASET TO UNDERSTAND HOW TO CALCULATE THE AVERAGE. 
  sum.salary.from = sum(NYC_Jobs[NYC_Jobs$`Civil Service Title` == i,"Salary Range From"])
  sum.salary.to = sum(NYC_Jobs[NYC_Jobs$`Civil Service Title` == i,"Salary Range To"])
  average.salary[index] = mean(sum.salary.from + sum.salary.to)
    
  # DO NOT TOUCH PLEASE
  index = index + 1
}

#Q4."Salary Breakdown".4 data frame creation
salary.dataframe <- data.frame(unique.agencies,offered.jobs,offered.values,average.salary)

#Q4."Salary Breakdown".5 ORDER THE DATASET BASED ON WHAT YOU WANT TO MEASURE
ordered.jobs <- salary.dataframe[order(-offered.values),]

#Q4."Salary Breakdown".6 Accessing the top 10 values in decreasing order

order.top10<- ordered.jobs[1:10,]

##### GRAPHING TIME

#Q4."Salary Breakdown".7.a.iii refer to the comment inside of the brackets below to help you through this division of values

x.input = unique(NYC_Jobs[which(NYC_Jobs$`Civil Service Title`== order.top10$offered.jobs),c("Agency","Salary Range From","Salary Range To")])$Agency



#################
#this is just for my refrence
# Different Approach as Sridhar guided me but x & y lengths are different . 
#for (i in 1:10){
#  x.input[i]<- NYC_Jobs[which(NYC_Jobs$`Civil Service Title` == order.top10[i,]$offered.jobs),]$Agency
#
#Q4."Salary Breakdown".7.a.iv use the same which condition from the x.input value to find the average of the Salary Range To and Salary Range From values
#  y.input[i] <- (NYC_Jobs[which(NYC_Jobs$`Civil Service Title` == order.top10[i,]$offered.jobs),]$`Salary Range To` +  NYC_Jobs[which(NYC_Jobs$`Civil Service Title` == order.top10[i,]$offered.jobs),]$`Salary Range From`)/2
# i = i+1
#}


##################

#Q4."Salary Breakdown".7.a.iv use the same which condition from the x.input value to find the average of the Salary Range To and Salary Range From values


y.input = (unique(NYC_Jobs[which(NYC_Jobs$`Civil Service Title`== order.top10$offered.jobs),c("Agency","Salary Range From","Salary Range To")])$`Salary Range From` + unique(NYC_Jobs[which(NYC_Jobs$`Civil Service Title`== order.top10$offered.jobs),c("Agency","Salary Range From","Salary Range To")])$`Salary Range To`)/2

index  = 1


#Q4."Salary Breakdown".7.c 
 
test.scatter<- plot_ly(x = ~x.input, 
        y = ~y.input,
        name= x.input ,
        type= "scatter") %>%
  add_markers(x = x.input[i], y.input[i], name=order.top10$offered.jobs[i]) %>%
  layout(title = 'Salary breakdown For All Jobs ')


test.scatter


```


Row {data-height=400}
-------------------------------------
   
### Top 10 Chart

```{r}
#Q4."Salary Breakdown".8
library(gridExtra)
library(grid)


# I dont know why this table doesent show up in the R Markdown file but this same command works in R script and gets the output table.
# I have attached screenshots of the same.
grid.table(order.top10)
order.top10




```   
    
### Description
x-input -> Agencies that have their Civil services title in top 10 
y-Input -> Average salaries of these Agencies.

the Scatter plot consist of the xinput vs yinput showing the average salaries of top 10 civil service titles with respective Agencies. 
Department of Envoirnment Proctection has highest average salary of around 140k